---
title: "Describe CWL Tools/Workflows in R and Execute it on Seven Bridges Supported Platform"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
    highlight: haddock
    css: style.css
---

<!--
%\VignetteIndexEntry{How to Describe CWL Tools/Workflows in R and Excute it on Seven Bridges Supported Platform}
%\VignettePackage{sevenbridges}
%\VignetteEngine{knitr::rmarkdown}
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown(css.files = "custom.css")
```

<!-- override white background for classless code chunks -->
<style type="text/css" scoped>
  pre:not([class]) { background-color: #F8F8F8; }
  pre code { background-color: transparent; }
</style>

```{r code, echo = FALSE}
code <- function(...) {
    cat(paste(..., sep = "\n"))
}

code2 <- function(...) {
    cat(paste("```markdown", ..., "\n", "```", sep = "\n"))
}
```

```{r global_options, include=FALSE}
library(BiocStyle)
knitr::opts_chunk$set(eval = TRUE)
```

# Describe Tools in R

The main interface provided by `sevenbridges` package is `Tool` function, it's basically a R interface similar to Seven Bridges's graphic user interface to describe tools, which I also highly recommend because it's very easy. 

For R users who want to script in everything here is alternative solution. I also did some work trying to make it simpler to use, any suggestions are welcomed. 

So I highly recommend user go over documentation [The Tool Editor](http://docs.cancergenomicscloud.org/docs/the-tool-editor) chapter for cancer genomic cloud to understand how it works, and even try it on the platform with the GUI. This will help use our R interface better and easier. 

Or you just want to create tools quickly, please keep reading. 


## Example: give me some randome number

Let's start from the most simple case

- Use exiting docker container
- directly execute from command line using `Rscript`

 

Some basic arguments used in `Tool` function.

- baseCommand: Specifies the program to execute. 
- stdout: Capture the command's standard output stream to a file written to the designated output directory. 
- inputs: inputs for your command line
- outputs: outputs you want to collect

For more accepted parameter, please check `help(CommandLineTool)` and `help(Tool)` for more details. Again we basically conform to CWL standard, you can always check their manual.

### Requirement and hints

In short, hints are not _required_ for execution. We now accept following requirement items `cpu`, `mem`, `docker`, `fileDef`; and you can easily construct them via `requirements()` constructor. 

The downside of this is clear, you have much less control over your script and command line interface. 

### Specify inputs and outpus

Most likely your command line interface accept extra arguments, for example, 

- file
- string
- enum
- int
- float
- boolean

So to specify that in your tool, you can use `input` function, then pass it to the `inputs` arguments as a list or single item. You can even construct them as data.frame, but in that way, you usually have to provide equal arguments, less typing but less flexible. 

`input()`  require arguments `id` and `type`. `output()`  require arguments `id` because `type` by default is file. 

### Using existing docker image and command

If you already have a docker image in mind that provide the functionality you need, you can just use it. The `baseCommand` is the command line you want to execute in that container. `stdout` specify the output file you want to capture the standard output and collect it on the platform. 

```{r}
library(sevenbridges)
rbx <- Tool(id = "runif", 
            label = "runif",
            hints = requirements(docker(pull = "rocker/r-base"), 
                                 cpu(1), mem(2000)), 
            baseCommand = "Rscript -e 'runif(100)'", 
            stdout = "output.txt",
            outputs = output(id = "random", glob = "*.txt"))

rbx
rbx$toJSON()
```

By default the tool object shows YAML, but you can simply convert it to JSON and copy it to your seven bridges platform graphic editor by importing JSON. 

```{r}
rbx$toJSON()
rbx$toJSON(pretty = TRUE)
rbx$toYAML()
```

### Add-on script

If you want to create simple script based on existing image, use `fileDef`. 

```{r}
## Make a new file
fd <- fileDef(name = "runif.R",
              content = "sed.seed(1)
                   runif(100)")

rbx <- Tool(id = "runif", 
            label = "runif",
            hints = requirements(docker(pull = "rocker/r-base"), 
                                 cpu(1), mem(2000)),
            requirements = requirements(fd),
            baseCommand = "Rscript runif.R",
            stdout = "output.txt",
            outputs = output(id = "random", glob = "*.txt"))   
rbx
```




```{r}
## pass a input list
in.lst <- list(input(id = "number",
                     description = "number of observations",
                     type = "integer",
                     label = "number",
                     prefix = "--n",
                     default = 1,
                     required = TRUE, 
                     cmdInclude = TRUE),
               input(id = "min",
                     description = "lower limits of the distribution",
                     type = "float",
                     label = "min",
                     prefix = "--min",
                     default = 0),
               input(id = "max",
                     description = "upper limits of the distribution",
                     type = "float",
                     label = "max",
                     prefix = "--max",
                     default = 1),
               input(id = "seed",
                     description = "seed with set.seed",
                     type = "float",
                     label = "seed",
                     prefix = "--seed",
                     default = 1))


## the same method for outputs
out.lst <- list(output(id = "random",
                       type = "file",
                       label = "output", 
                       description = "random number file",
                       glob = "*.txt"),
                output(id = "report",
                       type = "file",
                       label = "report", 
                       glob = "*.html"))


rbx <- Tool(id = "runif",
            label = "Random number generator",
            hints = requirements(docker(pull = "tengfei/runif"), 
                                 cpu(1), mem(2000)),
            baseCommand = "runif.R",
            inputs = in.lst, ## or ins.df
            outputs = out.lst)
```

Here I use data.frame as example for input and output. 

```{r}
in.df <- data.frame(id = c("number", "min", "max", "seed"),
                    description = c("number of observation", 
                                    "lower limits of the distribution",
                                    "upper limits of the distribution",
                                    "seed with set.seed"),
                    type = c("integer", "float", "float", "float"),
                    label = c("number" ,"min", "max", "seed"), 
                    prefix = c("--n", "--min", "--max", "--seed"),
                    default = c(1, 0, 10, 123), 
                    required = c(TRUE, FALSE, FALSE, FALSE))

out.df <- data.frame(id = c("random", "report"),
                     type = c("file", "file"),
                     glob = c("*.txt", "*.html"))

rbx <- Tool(id = "runif",
            label = "Random number generator",
            hints = requirements(docker(pull = "tengfei/runif"), 
                                 cpu(1), mem(2000)),
            baseCommand = "runif.R",
            inputs = in.df, ## or ins.df
            outputs = out.df)


```

### Execute the tool in the cloud

With API function, you can directly load your Tool into the account. Run a task, for "how-to", please check the API complete guide 

Following section, please for now skip. 

```{r, eval = FALSE}
a <- Auth(platform = "cgc", username = "tengfei")
p <- a$project("demo")
app.runif <- p$app_add("runif555", rbx)
aid <- app.runif$id
p$task_add(name = "Draft runif simple", 
           description = "Description for runif", 
           app = aid,
           inputs = list(min = 1, max = 10))
## confirm, show all task status is draft
(tsk <- p$task(status = "draft"))
tsk$run()
tsk$download("~/Downloads/")
```


# Wokrflow

__Graphic User Interface on Seven Bridges Platform is way more conventient__

## Introduction

To create a workflow, we provide simple interface to pipe your tool into a single workflow, it works under situation like

- Simple linear tool connection
- Output and input match at lest have one match, then we will connect it automatically.


__Note__ for complicated workflow construction, I highly recommend using our graphical interface to do it, there is no better way. 

## Example

Here I will give a quick example

1. Tool 1 output 1000 random number
2. Tool 2 take log on it 
3. Tool 3 do a mean calculation of everything

Here are methods we support

- Maximize the flexibility please use `Flow` constructor to construct your flow, then pass steps using `+` sign to connect your object. `+` only outputs `StepList` and support operation
    - Tool + Tool 
    - StepList + Tool
    - StepList + StepList
- Alternatively `%>%` will output `Flow` not `StepList`, if you connect tools with `%>%` we will create id for you. it always output `Workflow`
    - Tool %>% Tool 
    - Workflow %>% Tool
    - Workflow %>% Tool
    
### Construct tools first    

```{r}
library(sevenbridges)
## A tool that generate a 100 random number
t1 <- Tool(id = "runif new test 3", label = "random number",
           hints = requirements(docker(pull = "rocker/r-base")),
           baseCommand = "Rscript -e 'x = runif(100); write.csv(x, file = 'random.txt', row.names = FALSE)'", 
           outputs = output(id = "random", 
                            type = "file", 
                            glob = "random.txt"))

## A tool that take log
fd <- fileDef(name = "log.R",
              content = "args = commandArgs(TRUE)
                         x = read.table(args[1], header = TRUE)[,'x']
                         x = log(x)
                         write.csv(x, file = 'random_log.txt', row.names = FALSE)
                         ")

t2 <- Tool(id = "log new test 3", label = "get log",
           hints = requirements(docker(pull = "rocker/r-base")),
           requirements = requirements(fd),
           baseCommand = "Rscript log.R", 
           inputs = input(id = "number",
                           type = "file"),
           outputs = output(id = "log", 
                            type = "file", 
                            glob = "*.txt"))

## A tool that do a mean
fd <- fileDef(name = "mean.R",
              content = "args = commandArgs(TRUE)
                         x = read.table(args[1], header = TRUE)[,'x']
                         x = mean(x)
                         write.csv(x, file = 'random_mean.txt', row.names = FALSE)
                          ")

t3 <- Tool(id = "mean new test 3", label = "get mean",
           hints = requirements(docker(pull = "rocker/r-base")),
           requirements = requirements(fd),
           baseCommand = "Rscript mean.R", 
           inputs = input(id = "number",
                           type = "file"),
           outputs = output(id = "mean", 
                            type = "file", 
                            glob = "*.txt"))

steplist <- t1 + t2 + t3
steplist
```


### Connect tools to a flow

To create a Flow we suggest you using `Flow` function, so that you can pass id and label to it. 

```{r}
f <- Flow(id = "Random-log-mean-new-test-2", 
          label = "random log mean new test", 
          steps = steplist)
f$toJSON()
f$toJSON(pretty = TRUE)
```

Or use %>% 

```{r}
f <- t1 %>% t2 %>% t3
f$toJSON()
```

If it's the first time you upload those tools, following script will push both 
tools and the workflow to your project.

```{r, eval = FALSE}
## need id, full, sbg:id
library(sevenbridges)
a <- Auth(platform = "cgc", username = "tengfei")
p <- a$project(id = "tengfei/helloworld")
app.runif <- p$app_add("new_flow", f)
aid <- app.runif$id
p$task_add(name = "Draft flow", 
           description = "Flow test", 
           app = aid)
## confirm, show all task status is draft
(tsk <- p$task('Draft flow'))
tsk$run()
```

You can get App from your repos and connect it in R to build flow.

```{r, eval = FALSE}
app1 <- p$app(id = "tengfei/helloworld/runif_new_test/0")
app2 <- p$app(id = "tengfei/quickstart/log_new_test_3/0")
app3 <- p$app(id = "tengfei/quickstart/log_new_test_3/0")

f <- app1 %>% app2 %>% app3

f$id <- "randome number test flow 2"
f$"sbg:id" <- "tengfei/helloworld/testtesttest"
f$toJSON(pretty = TRUE)
```

** Important ** 

This flow construction method is easy for linear flow, your best choice is always our graphic user interface and editor. 


