---
title: "Tutorials: Creating Your First Docker App with Command Line Interfqce and Report."
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
    highlight: haddock
    css: style.css
---

<!--
%\VignetteIndexEntry{Tutorials: Creating Your First Docker App and Command Line Interface and Report.}
%\VignettePackage{sevenbridges}
%\VignetteEngine{knitr::rmarkdown}
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown(css.files = "custom.css")
```

<!-- override white background for classless code chunks -->
<style type="text/css" scoped>
  pre:not([class]) { background-color: #F8F8F8; }
  pre code { background-color: transparent; }
</style>

```{r code, echo = FALSE}
code <- function(...) {
    cat(paste(..., sep = "\n"))
}

code2 <- function(...) {
    cat(paste("```markdown", ..., "\n", "```", sep = "\n"))
}
```

```{r global_options, include=FALSE}
library(BiocStyle)
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction


In this tutorial, we will go through couple different ways to 

- Creating Your CWL App
- Make native command line interface in R

There are many fun ways to do it, here I am more focused for R developers. 

I will start from the easiest one to most formal one. 

- When you have a docker image ready, just need to describe a tool
- You want to create your own docker container with command line interface

## Existing docker images

Usually it's recommended you may want to always start with a 'core' image in your group
or communty to serve as the basis for your tool chain. 

Fortunately, for R developers, there are official R and Biocondcutor docker images already
I will introduce briefly in following section.

### Rocker Project

Official R Docker images is called "Rocker"" project and is on [github](https://github.com/rocker-org/rocker), please visit the 
page to find more details and Dockerfile.

| Image         | Description          | 
| ------------- |:-------------:|
| rocker/r-base     | base package to build from| 
| rocker/r-devel     | base plus R-devel from SVN |  
| rocker/rstudio | base plus RStudio Server  |    
| rocker/hadleyverse | rstudio + Hadley's packages, LaTeX |
| rocker/ropensci | hadleyverse + rOpenSci packages |
| rocker/r-devel-san | base, SVN's R-devel and SAN |


### Bioconductor Images

Bioconductor have a nice [page](http://bioconductor.org/help/docker/) about the official docker images, please read for more details.

| Image | 
| ---- | 
| bioconductor/devel_base
| bioconductor/devel_core
| bioconductor/devel_flow |
| bioconductor/devel_microarray |
| bioconductor/devel_proteomics |
| bioconductor/devel_sequencing |
| bioconductor/release_base |
| bioconductor/release_core |
| bioconductor/release_flow |
| bioconductor/release_microarray |
| bioconductor/release_proteomics |
| bioconductor/release_sequencing |

To understand the image quickly here is the short instruction for the image name:

- __release__ images are based on `rocker/rstudio`
- __devel__ images are based on `rocker/rstudio-daily`
- __base__: Contains R, RStudio, and BiocInstaller + system dependencies.
- __core__: base + a selection of core.
- __flow__: core + all packages tagged with the _FlowCytometry_ biocView.
- __microarray__: core + all packages tagged with the _Microarray_ biocView.
- __proteomics__: core + all packages tagged with the _Proteomics_ biocView.
- __sequencing__: core + all packages tagged with the _Sequencing_ biocView.

### Dockerhub 

[Dockerhub](https://hub.docker.com/) also provide public/private repos, you can search
existing tools without building yourself, it's very likely some popular tool already have
docker container well maintained there.

### Seven Bridges Docker Registry

If you plan to develop/execute/deploy on any of Seven Bridges Platforms, including cancer genomics cloud. Users are encouraged to use Seven Bridges Registry, so the push and pulling with be faster. 

# Tutorial: random number generator 

Ok not "hello world"" example this time : ) will leave that as exercise to users, this time I  will start with another simple example for R users, but with more arguments to play with. 

Our goal here is to making a cwl app to generate unif random numbers, yes, the core
function is `runif()`, it's native function in R. 

```{r}
runif
runif(10)
```

## Using exiting images

## Building your own contains and app

### Using docopt package

In R, we also have a nice implementation in a package called `docopt`, developed by *Edwin de Jonge*. Check out its [tutorial](https://github.com/docopt/docopt.R) on github.

So let's quickly create a command line interface for our R scripts with a dummy example. Let's turn the uniform distribution function `runif` into a command line tool.

when you check out the help page for `runif`, here is the key information you want to mark down. 

```
Usage

runif(n, min = 0, max = 1)

Arguments

n	
number of observations. If length(n) > 1, the length is taken to be the number required.

min, max	
lower and upper limits of the distribution. Must be finite.
```

I will add one more parameter to set seed, here is the R script file called `runif.R`. 

At the beginning, I use docopt standard to write my tool help. 
```{r}
'usage: runif.R [--n=<int> --min=<float> --max=<float> --seed=<float>]

options:
 --n=<int>        number of observations. If length(n) > 1, the length is taken to be the number required [default: 1].
 --min=<float>   lower limits of the distribution. Must be finite [default: 0].
 --max=<float>   upper limits of the distribution. Must be finite [default: 1].
 --seed=<float>  seed for set.seed() function [default: 1]' -> doc

library(docopt)

```
Let's first do some testing in R session before you make it a full functional command line tool.

```{r}
docopt(doc) #with no argumetns provided
docopt(doc, "--n 10 --min=3 --max=5")
```

Add my command line function

```{r}
opts <- docopt(doc)
set.seed(opts$seed)
runif(n = as.integer(opts$n), 
      min = as.numeric(opts$min), 
      max = as.numeric(opts$max))
```

Add Shebang at the top of the file, and a complete example for `runif.R` command line will be like this

```{r, eval=FALSE}
#!/usr/bin/Rscript
'usage: runif.R [--n=<int> --min=<float> --max=<float> --seed=<float>]

options:
 --n=<int>        number of observations. If length(n) > 1, the length is taken to be the number required [default: 1].
 --min=<float>   lower limits of the distribution. Must be finite [default: 0].
 --max=<float>   upper limits of the distribution. Must be finite [default: 1].
 --seed=<float>  seed for set.seed() function [default: 1]' -> doc

library(docopt)
opts <- docopt(doc)
set.seed(opts$seed)
runif(n = as.integer(opts$n), 
      min = as.numeric(opts$min), 
      max = as.numeric(opts$max))
```

OK seems good, now let's test it in our terminal, don't forget to make it executable by doing something like `chmod 755 runif.R`

```
$ ./runif.R --help
Loading required package: methods
usage: runif.R [--n=<int> --min=<float> --max=<float> --seed=<float>]

options:
 --n=<int>        number of observations. If length(n) > 1, the length is taken to be the number required [default: 1].
 --min=<float>   lower limits of the distribution. Must be finite [default: 0].
 --max=<float>   upper limits of the distribution. Must be finite [default: 1].
 --seed=<float>  seed for set.seed() function [default: 1]
$ ./runif.R
Loading required package: methods
[1] 0.2655087
$ ./runif.R
Loading required package: methods
[1] 0.2655087
$ ./runif.R --seed=123 --n 10 --min=1 --max=100
Loading required package: methods
 [1] 29.470174 79.042208 41.488715 88.418723 94.106261  5.510093 53.282443
 [8] 89.349485 55.592066 46.204859
```

# Exercise

I leave this as excercise to users, please have fun. anwsers will be added later. Please email me your solution or share with us. 

## Produce graphics

- Try to use default graphic device
- Try ggplot2 to generate a png file and capture it in output

## Produce report 

- Try produce a report based on your knitr template

## Deploy shiny apps

- Try to deploy a shiny report on shinyapps.io
