---
title: "Running sevenbridges SDK image and launch Rstudio server"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
    highlight: haddock
    css: style.css
---

<!--
%\VignetteIndexEntry{Running sevenbridges SDK image and launch Rstudio server}
%\VignettePackage{sevenbridges}
%\VignetteEngine{knitr::rmarkdown}
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown(css.files = "custom.css")
```

<!-- override white background for classless code chunks -->
<style type="text/css" scoped>
  pre:not([class]) { background-color: #F8F8F8; }
  pre code { background-color: transparent; }
</style>

```{r code, echo = FALSE}
code <- function(...) {
    cat(paste(..., sep = "\n"))
}

code2 <- function(...) {
    cat(paste("```markdown", ..., "\n", "```", sep = "\n"))
}
```

```{r global_options, include=FALSE}
library(BiocStyle)
knitr::opts_chunk$set(eval = TRUE)
```

# Introduction

The goal of this docker image is to provide

- SDK environment that includes
    - Rabix python package
    - Command line uploader
    - Seven Bridgs Command line tools
    - sevenbridges R package
    - In API you can call uploader
- Rstudio: you can launch the server, I will show you how later in this tutorial

## Dockerfile

The Dockerfile is included with the package in `inst/docker` folder. 

```{r}
fl <- system.file("docker/sevenbridges/devel", "Dockerfile", package = "sevenbridges")
fl
```
Here is the current content of Dockerfile
```{r comment=''}
cat(readLines(fl), sep = '\n')
```

# Docker hub autobuild

A hook is added to build the docker image automatically from my [github](https://github.com/tengfei/sevenbridges), it's under `inst/docker`, it's automatically build on my dockerhub [page](https://hub.docker.com/r/tengfei/sevenbridges/). you can directly use this `tengfei/sevenbridges`

# Launch Rstudio server 

For example, you can ssh into your AWS instance, here I suppose you already have docker installed,  and pull the image 

```
docker pull tengfei/sevenbridges
```

To launch sevenbridges Rstudio server image, I recommend you read this [tutorial](https://github.com/rocker-org/rocker/wiki/Using-the-RStudio-image)

Or following the quick instruction here

```
docker run -d -p 8787:8787 tengfei/sevenbridges
docker run -d -p 8787:8787 -e USER=<username> -e PASSWORD=<password> rocker/rstudio
```

You will be able to access the Rstudio in the browser by something like

`http://<your ip address>:8787`

Sometimes you want to add more users, to add users

```
## Enter the container
docker exec -it <container-id> bash
## Interactively input password and everything else
adduser <username>
```


# Shiny Apps

## User ShinyApp app



## Add hook on the Rstudio server

```

```


